---
title: "Reduced-rank models in glmmTMB"
output: html_document
---

### House work 

First call libraries
```{r message=FALSE, warning=FALSE}

packages <- c("ggplot2", "dplyr", "tidyr", "glmmTMB", "ecostats", "DHARMa")
install.packages(setdiff(packages, rownames(installed.packages()))) 
invisible(lapply(packages, library, character.only = TRUE))

source("functions.R")
```

### Wind farm example

The wind farm data set is part of the `ecostats` package. It is in wide format with Species along the columns. We need to transform the data to long format for `glmmTMB`

```{r}
wf <- cbind(windFarms$abund, windFarms$X) %>% 
  mutate(ID = 1:nrow(.)) %>% 
  pivot_longer(cols = colnames(windFarms$abund), 
               names_to = "Species",
               values_to = "Abundance")
```

As always, we need to "clean" the data.
```{r}
stat03 <- unique(wf[wf$Year=="2003",]$Station)
stat10 <- unique(wf[wf$Year=="2010",]$Station)
sum_abund <- wf %>%
  filter(Station %in% stat03 ) %>%
  group_by(Species) %>%
  summarise(obs = sum(Abundance > 0) ) %>% 
  ungroup() %>% 
  arrange(-obs) # order from most to least abundant

spp_names <- unique(sum_abund$Species[sum_abund$obs>=5])
wf_data <- wf %>% 
  filter(Species %in% spp_names) %>% 
  mutate(Species = factor(Species, levels = spp_names))
```

## Plot the data

```{r}
labs=c(0,1,10,100)
wf_boxplot <- wf_data %>% 
  ggplot(aes(x = Year, y = log(Abundance + 1), fill = Zone)) +
  geom_boxplot(outlier.shape = 21)+
  facet_wrap(~ Species, ncol = 3, scales = "free") +
  labs(x = "Year", color  = "Zone", y = "Abundance [log(y+1)-scale]") +
  scale_y_continuous( breaks = log(labs + 1), labels = labs)+
  theme_classic() +
  theme(legend.position="bottom"  ) +
  scale_fill_brewer(palette = "Set2")

wf_boxplot
```

## Fitting mixed models in glmmTMB

`glmmTMB` uses the same formula syntax as `lme4`.
```{r, eval = F}
?glmmTMB
```

For the wind farm example, we would like to fit this GLMM:

```{r}
fit.us <- glmmTMB(Abundance ~ Year*Zone + diag(Year*Zone|Species) + 
                 (1 | Station) +
                 (Species + 0|ID),
               wf_data, family = poisson)
```

## reduced-rank models

Let's try to fit a reduced-rank model

```{r}

```
